\section{The Digital Image}

\begin{definition}[Problems]
   Transmission inference, artifacts, scratches, noise, bad contrast / resolution, motion blur, spilling
\end{definition}

\begin{definition}[Image]
  Pattern of a value varying in space and/or time.
\end{definition}

\begin{definition}[Pixel]
  Discrete samples of a continuous image function.
\end{definition}

\subsection{Sensors}
\begin{definition}[Charge Coupled Device (CCD)]
  Contains sensor array (array of photosites [bucket of electrical charge]), ADC (digitalize measurement line by line) and image array. \textit{Sequential} readout.
  \begin{itemize}[label=-]
    \item \textbf{Blooming}: Oversaturation of photosites \(\implies\) flooding.
    \item \textbf{Bleeding/Smearing}: Accumulate charges during readout.
    \item \textbf{Dark Current}: Thermally generated charge \(\Rightarrow\) non-0 output. Reduced by cooling or avg. dark images and subtract.
  \end{itemize}
\end{definition}

\begin{definition}[CMOS]
  CCD, but with amplifiers.
  \begin{itemize*}
    \item cheaper
    \item lower power
    \item per pixel amplification
    \item random pixel access
    \item no blooming
    \item on chip integration
    \item rolling shutter
    \item noise
    \item less sensitive
  \end{itemize*}

  For videos: rolling shutter due to sequential read-out of lines.
\end{definition}

\begin{definition}[Dynamic Vision Sensor (DVS)]
  Smart pixels: ind. and async.
\end{definition}

\subsection{Sampling and Quantization}
\begin{definition}[Sampling]
  \begin{itemize*}
    \item Cartesian (grid)
    \item Hexagonal
    \item non-uniform
  \end{itemize*}

  Reconstruction with e.g. bilinear interpolation.
\end{definition}

\begin{definition}[Quantization]
  \(\R \to \Z\), lossy, simple quant. with \(k = 2^b\) levels.
\end{definition}

\begin{definition}[Resolutions]
  \begin{itemize*}
    \item \textit{Image}: px \(\times\) px
    \item \textit{Geometric}: \#pixels per area
    \item \textit{Radiometric}: \#bits per pixel
  \end{itemize*}
\end{definition}

\begin{definition}[Additive Gaussian Noise]
  \(I(x, y) = f(x, y) + c, \ c \sim \normal(0, \sigma^2)\)
\end{definition}

\begin{definition}[Poisson Noise]
  Shot noise for low light as \(\normal\) can be \(< 0\).
\end{definition}

\begin{definition}[Quantization Error]
  \(c \sim \unif([-0.5, 0.5], [-0.5, 0.5])\)
\end{definition}

\begin{definition}[Rician noise]
  Noise that appears in MRI.
\end{definition}

\begin{definition}[SNR]
  Index of quality: \(s = \frac{F}{\sigma}, \quad F = \frac{1}{XY}\sum_{x=1}^X\sum_{y=1}^Y f(x, y)\). \textit{PSNR} (Peak SNR) \(s_\text{peak} = \frac{F_{\max}}{\sigma}\) often used in practice.
\end{definition}

\subsection{Color Cameras}
\begin{definition}[Prism]
  Separate light in 3 beams to 3 sensors.
\end{definition}

\begin{definition}[Filter Mosaic]
  Coat rasterized filter (filter out different colors).
\end{definition}

\begin{definition}[Filter Wheel]
  Rotate multiple filters in front of lens.
\end{definition}

\begin{definition}[CMOS]
  Absorb colors at different depths \(\implies\) better quality.
\end{definition}

\begin{tabularx}{\linewidth}{|X|X|X|X|}
  \hline
  Approach & Prism & Mosaic & Wheel \\ \hline
  \# Sensors & 3 & 1 & 1 \\ \hline
  Separation & High & Avg. & Good \\ \hline
  Cost & High & Low & Average \\ \hline
  Framerate & High & High & Low \\ \hline
  Artifacts & Low & Aliasing & Motion \\ \hline
  Bands & 3 & 3 & \(\geq 3\) \\ \hline
  Usage & High-End & Low-end & Scientific \\ \hline
\end{tabularx}
