\section{Egocentric Vision}
\begin{definition}[FPV]
    \adv{Dynamic uncurated content, long-form video stream, driven by goals, interactions and attention}.
    \disadv{occlusion, motion blur}.
\end{definition}

\begin{definition}[TPV]
    \adv{static setup, controlled field of view}
    \disadv{curated moment in time, disembodied}.
\end{definition}

\begin{definition}[Gaze Tracker]
    Eye movements and track correlation to actions.
\end{definition}

\begin{definition}[Ego4D Benchmarks and Challenges]
    \begin{itemize*}
        \item \textit{Past}: episodic memory (where is \(X\))
        \item \textit{Present}: Hands \& Objects: what am I doing? who said what? who is attending?
        \item \textit{Future}: Forecasting
    \end{itemize*}
\end{definition}

\begin{definition}[Egocentric Vision Tasks]
    \begin{itemize*}
        \item Localisation
        \item 3D scene understanding
        \item Anticipation
        \item Action Recognition
        \item Gaze Understanding and Prediction
        \item Social behaviour understanding
        \item full body pose estimation
        \item hand and hand-object interactions
        \item person identification
        \item privacy
        \item summarisation
        \item visual question answering
    \end{itemize*}
\end{definition}

\begin{definition}[Hand Interaction]
    Use \textit{Faster-RCNN}, predict bounding box, left/right side, contact state (none/self/person/portable/non-portable object)
\end{definition}

\begin{definition}[HaMeR]
    Output 3D Mesh from monocular image,s left/right hand info using ViT.
\end{definition}

\begin{definition}[TransFusion]
    Use past-observable frames + predition frame for LLM features. LMM features + predition frame go to Multimodal Fusion Model and predicts location + (VERB, NOUN, TTC).
\end{definition}

\begin{definition}[Language Context Generation]
    Get (verb, noun) pairs using CLIP and aggregate action context.
\end{definition}

\begin{definition}[PALM]
    Predict actions through language models.
\end{definition}

\begin{definition}[Egocentric Gaze Estimation]
    Using Path Complexity Index (PCI), prediction for example car direction using gaze.
\end{definition}

\begin{definition}[LMK]
    Lift (2D obs. to 3D coords), Match (connect obs over time), Keep (maintain object track).
\end{definition}
