\section{Accelerating Data Structures}

\begin{definition}[Few intersection computations]
  \begin{itemize*}
    \item uniform grid
    \item binary space partition tree, kd tree, octree
    \item bounding volume hierarchies
  \end{itemize*}
\end{definition}

\begin{definition}[Fewer rays]
  \begin{itemize*}
    \item early ray termination
    \item adaptive sampling
  \end{itemize*}
\end{definition}

\begin{definition}[Generalized rays]
  \begin{itemize*}
    \item beam tracing
    \item cone tracing
  \end{itemize*}
\end{definition}

\begin{definition}[Ray Tracing Cost]
  \(\O(\#\text{px} \cdot \#\text{Objs})\), thus for each ray \(\O(\#\text{Objs})\)
\end{definition}

\subsection{Spatial Decomposition}
Insert objects into disjoint spatial regions. Top-Down.

\begin{definition}[Axis Aligned Bounding Boxes (AABB)]
  Each box is defined by two vectors. Easy check if intersection into box.
\end{definition}

\begin{definition}[Uniform Grid]
  Partition space into voxels.
  Each cell contains primitives that overlap cell.
  Resolution heuristic: \(3 \sqrt[3]{n}\).
  Store reference for each object in cell.
  Incrementally rasterize ray.

  \textbf{Problems}: non-uniform scenes (use hierarchical grids)
\end{definition}

\begin{definition}[Octree]
  Adaptive binary splits like uniform grid. Recursively subdivide each cell into 8 \textbf{equal} subcells.
\end{definition}

\begin{definition}[KD-Tree]
  Recursively binary split via \textbf{axis-aligned} planes until termination criteria (number of objects or maximum depth). Only leaf nodes store reference to geometry.
\end{definition}

\begin{definition}[BSP-Trees]
  Recursively split space using \textbf{arbitrary} planes.
\end{definition}

\begin{theorem}
  Complexity increases from Octree to KD to BSP, but it increases the flexibility and has fewer intersections.
\end{theorem}

\subsection{Object decomposition}
Partition objects into disjoint sets where the bounding volumes may overlap.
Either bottom-up or top-down.

\begin{definition}[Bounding Volumes]
  \begin{itemize*}
    \item Spheres
    \item AABB
    \item Oriented bounding box (OBB)
    \item K-discrete orientation polytopes
    \item Convex hulls
  \end{itemize*}

  \textit{Tradeoff}: complex, tight fit, fewer intersections \(\leftrightarrow\) simple, fast
\end{definition}