\section{Visibility and Shadows}

\begin{algorithm}[Painter's algorithm]
  Render from furthest to nearest.

  \textit{Problems}:
  \begin{itemize*}
    \item cyclic overlaps
    \item intersections
  \end{itemize*}
\end{algorithm}

\begin{algorithm}[Z-Buffering]
  Store depth to nearest object per pixel.

  \begin{enumerate}
    \item Initialize all \(z\) values to \(\infty\)
    \item For each polygon: If new value smaller, replace!
  \end{enumerate}

  \textit{Problems}:
  \begin{itemize*}
    \item limited resolution
    \item resolution is non-linear
  \end{itemize*}
\end{algorithm}

\subsection{Shadows}

\begin{definition}[Planar Shadows]
  Draw projection on the ground.

  \textit{Limitations}:
  No self shadows or on other objects.
  Problems with curved surfaces.
\end{definition}

\begin{definition}[Projective texture shadows]
  Separate obstacle and receiver. Compute b/w image from light and use as projective texture.

  \textit{Limitations}:
  Need to specify obstacle \& receiver.
  No self-shadows.
\end{definition}

\begin{algorithm}[Shadow Maps]
  Compute depths from light \(d(x_L)\) and camera.
  For each pixel in camera plane:
  \begin{enumerate}
    \item Compute point in world coordinates
    \item Project onto light plane \(z_L\)
    \item If \(d(x_L) < z_L\), then \(x\) is in shadow.
  \end{enumerate}

  Add bias for stability (\(d(x_L) + b < z_L\)).

  A point to shadow can be outside the FoV of shadow map, thus use cubical shadow map or spot lights.

  Should \textbf{not} filter depth, but take weighted average.
\end{algorithm}

\begin{algorithm}[Shadow Volumes]
  Explicitly represent the volume of space in shadow. If polygon in volume, it is in shadow.

  \begin{enumerate}
    \item Shoot a ray from the camera
    \item ++/-- counter each time volume is intersected.
    \item if counter \(> 0\), then primitive is in shadow
  \end{enumerate}

  Use silhouette edges only!

  \textit{Limitations}:
  Lots of geometry.
  Expensive to rasterize long skinny triangles.
  Object must be watertight
  Rasterization of polygons sharing an edge must not overlap and not have gap.
\end{algorithm}
